One-armed spirals in locally isothermal, radially structured self-gravitating discs

### 2014-11-20 paper/appendix.tex
minor updates
spell checked

We begin with the linearised equation of motion in terms of the
Lagragian displacement $\bm{\xi}$ as given by \cite{lin93b} but with an
additional potential perturbation, 
\begin{align}\label{lagragian_pert}
  &\frac{D^2\bm{\xi}}{Dt^2} +
  2\Omega\hat{\bm{z}}\times\frac{D\bm{\xi}}{Dt}  \notag \\ &= -
  \frac{\nabla \delta p }{\rho} + \frac{\delta\rho}{\rho^2}\nabla\rho  
  -\nabla\delta\Phi_d - R
  \hat{\bm{R}}\left(\bm{\xi}\cdot\nabla\Omega^2\right) \notag \\
  & = -\nabla\left(\frac{\delta p}{\rho} + \delta \Phi_d\right) -
  \frac{\delta p}{\rho}\frac{\nabla\rho}{\rho} +
  \frac{\delta\rho}{\rho}\frac{\nabla p}{\rho} - - R
  \hat{\bm{R}}\left(\bm{\xi}\cdot\nabla\Omega^2\right),
\end{align}
where $D/Dt \equiv \p_t + \ii m \Omega$ for perturbations with
azimuthal dependence in the form $\exp\left(\ii m \phi\right)$, and  
the $\delta$ quantities denote Eulerian perturbations.

As explained in \cite{lin11b}, a conservation law for the angular
momentum of the perturbation may be obtained by taking the dot product
between Eq. \ref{lagragian_pert} and $(-m/2)\rho\bm{\xi}^*$, then
taking the imaginary part afterwards. The left hand side bcomes the
angluar momentum density. The first term on the right hand side (RHS)
becomes 
\begin{align}\label{angflux1}
  &-\frac{m}{2}\imag\left[-\rho\bm{\xi}^*\cdot\nabla\left(\frac{\delta p}{\rho} + \delta
    \Phi_d\right)\right] \notag\\ 
&= \frac{m}{2}\imag\left\{\nabla\cdot\left[\rho\bm{\xi}^*\left(\frac{\delta p}{\rho} + \delta
    \Phi_d \right) + \frac{1}{4\pi
    G}\delta\Phi_d\nabla\delta\Phi_d^*\right]\right\} \notag\\
&+ \frac{m}{2}\imag\left(\delta\rho^*\frac{\delta p}{\rho}\right),
\end{align}
where $\delta\rho = - \nabla\cdot\left(\rho\bm{\xi}\right)$ and
$\nabla^2\delta\Phi_d = 4\pi G \delta \rho$ have been used. The first
term on the RHS of Eq. \ref{angflux1} is (minus) the 
angular momentum flux. The second term on RHS of Eq. \ref{angflux1}, 
together with the remaining terms on the RHS of 
Eq. \ref{lagragian_pert} constitutes the background torque. That is, 
\begin{align}\label{tbg_3d}
  T_\mathrm{BG} = \frac{m}{2}\imag\left[
    \frac{\delta p}{\rho} \Delta\rho^* -
%\left(\delta\rho^* + \bm{\xi}^*\cdot\nabla\rho\right) - 
    \frac{\delta\rho}{\rho}\bm{\xi}^*\cdot\nabla p  
    + \rho \xi_R^*\xi_z\frac{\p\left(R\Omega^2\right)}{\p z}\right],
\end{align}
where $\Delta\rho = \delta\rho + \bm{\xi}\cdot\nabla\rho$ is the
Lagragian density perturbation.

@@

\section{Relation between horizontal Lagragian displacements for
  local, low frequency $m=1$ disturbances}\label{horizontal_displacements}
Here, we aim to relate the horizontal Lagrangian displacements $\xi_R$
and $\xi_\phi$ in the local approximation. Using the local solution to
the Poisson equation 
\begin{align}
  \delta \Phi_m = -\frac{2\pi G}{|k|} \delta\Sigma_m 
\end{align}
\citep{shu91}, the linearised azimuthal equation of motion becomes 
\begin{align} 
  - \ii\sbar \delta v_{\phi m}  + \frac{\kappa^2}{2\Omega}\delta v_{Rm} = -\frac{\ii
    m}{R\Sigma}\left(c_s^2 - \frac{2\pi G
      \Sigma}{|k|}\right)\delta\Sigma_m. 
\end{align}
Next, we replace the surface density perturbation
$\delta \Sigma_m = -\ii k \Sigma \xi_R$, and use the expressions
\begin{align}
  &\delta v_{Rm} = -\ii\sbar\xi_R,\\
  &\delta v_{\phi m} = -\ii\sbar\xi_\phi - \frac{\ii R
    \p_R\Omega}{\sbar} \delta v_{Rm}
\end{align}
\citep{papaloizou85} to obtain 
\begin{align}
  -\sbar^2\xi_\phi - 2\ii\sbar\Omega \xi_R =
  \frac{m}{kR}\left(\kappa^2 - \sbar^2\right)\xi_R, 
\end{align}
where the dispersion relation Eq. \ref{dispersion} was used. 
In the local approximation, $|kR|\gg m$ by assumption.%  Furthermore,  
% Also, for low frequency $m=1$ modes $\sbar \simeq -\Omega$, so the
% quantity $|\kappa^2 - \sbar^2| \ll \Omega^2$ in a nearly Keplerian disc
% ($\kappa\sim \Omega$). 
 Hence the RHS of this equation can be neglected. Then 
\begin{align}
  \xi_\phi \simeq -\frac{2\ii\Omega}{\sbar}\xi_R.
\end{align}
For low-frequency $m=1$ modes we have $\sbar \simeq -\Omega$, so
$\xi_\phi\simeq 2\ii\xi_R$, as used in the main text.

>>>>

We begin with the linearised equation of motion in terms of the
Lagrangian displacement $\bm{\xi}$ as given by \cite{lin93b} but with an
additional potential perturbation, 
\begin{align}\label{lagragian_pert}
  &\frac{D^2\bm{\xi}}{Dt^2} +
  2\Omega\hat{\bm{z}}\times\frac{D\bm{\xi}}{Dt}  \notag \\ &= -
  \frac{\nabla \delta p }{\rho} + \frac{\delta\rho}{\rho^2}\nabla\rho  
  -\nabla\delta\Phi_d - R
  \hat{\bm{R}}\left(\bm{\xi}\cdot\nabla\Omega^2\right) \notag \\
  & = -\nabla\left(\frac{\delta p}{\rho} + \delta \Phi_d\right) -
  \frac{\delta p}{\rho}\frac{\nabla\rho}{\rho} +
  \frac{\delta\rho}{\rho}\frac{\nabla p}{\rho} -  R
  \hat{\bm{R}}\left(\bm{\xi}\cdot\nabla\Omega^2\right),
\end{align}
where $D/Dt \equiv \p_t + \ii m \Omega$ for perturbations with
azimuthal dependence in the form $\exp\left(\ii m \phi\right)$, and  
the $\delta$ quantities denote Eulerian perturbations.

As explained in \cite{lin11b}, a conservation law for the angular
momentum of the perturbation may be obtained by taking the dot product
between Eq. \ref{lagragian_pert} and $(-m/2)\rho\bm{\xi}^*$, then
taking the imaginary part afterwards. The left hand side becomes the
angular momentum density. The first term on the right hand side (RHS)
becomes 
\begin{align}\label{angflux1}
  &-\frac{m}{2}\imag\left[-\rho\bm{\xi}^*\cdot\nabla\left(\frac{\delta p}{\rho} + \delta
    \Phi_d\right)\right] \notag\\ 
&= \frac{m}{2}\imag\left\{\nabla\cdot\left[\rho\bm{\xi}^*\left(\frac{\delta p}{\rho} + \delta
    \Phi_d \right) + \frac{1}{4\pi
    G}\delta\Phi_d\nabla\delta\Phi_d^*\right]\right\} \notag\\
&+ \frac{m}{2}\imag\left(\delta\rho^*\frac{\delta p}{\rho}\right),
\end{align}
where $\delta\rho = - \nabla\cdot\left(\rho\bm{\xi}\right)$ and
$\nabla^2\delta\Phi_d = 4\pi G \delta \rho$ have been used. The 
terms in square brackets on the RHS of Eq. \ref{angflux1} is (minus) the 
angular momentum flux. The second term on RHS of Eq. \ref{angflux1}, 
together with the remaining terms on the RHS of 
Eq. \ref{lagragian_pert} constitutes the background torque. That is, 
\begin{align}\label{tbg_3d}
  T_\mathrm{BG} = \frac{m}{2}\imag\left[
    \frac{\delta p}{\rho} \Delta\rho^* -
%\left(\delta\rho^* + \bm{\xi}^*\cdot\nabla\rho\right) - 
    \frac{\delta\rho}{\rho}\bm{\xi}^*\cdot\nabla p  
    + \rho \xi_R^*\xi_z\frac{\p\left(R\Omega^2\right)}{\p z}\right],
\end{align}
where $\Delta\rho = \delta\rho + \bm{\xi}\cdot\nabla\rho$ is the
Lagrangian density perturbation.

@@

\section{Relation between horizontal Lagrangian displacements for
  local, low frequency \lowercase{$m=1$} disturbances}\label{horizontal_displacements}
Here, we aim to relate the horizontal Lagrangian displacements $\xi_R$
and $\xi_\phi$ in the local approximation. Using the local solution to
the Poisson equation 
\begin{align}
  \delta \Phi_m = -\frac{2\pi G}{|k|} \delta\Sigma_m 
\end{align}
\citep{shu91}, the linearised azimuthal equation of motion becomes 
\begin{align} 
  - \ii\sbar \delta v_{\phi m}  + \frac{\kappa^2}{2\Omega}\delta v_{Rm} = -\frac{\ii
    m}{R\Sigma}\left(c_s^2 - \frac{2\pi G
      \Sigma}{|k|}\right)\delta\Sigma_m. 
\end{align}
Next, we replace the surface density perturbation
$\delta \Sigma_m = -\ii k \Sigma \xi_R$, and use the expressions
\begin{align}
  &\delta v_{Rm} = -\ii\sbar\xi_R,\\
  &\delta v_{\phi m} = -\ii\sbar\xi_\phi - \frac{\ii R
    \p_R\Omega}{\sbar} \delta v_{Rm}
\end{align}
\citep{papaloizou85} to obtain 
\begin{align}
  -\sbar^2\xi_\phi - 2\ii\sbar\Omega \xi_R =
  \frac{m}{kR}\left(\kappa^2 - \sbar^2\right)\xi_R, 
\end{align}
where the dispersion relation Eq. \ref{dispersion} was used. 
In the local approximation, $|kR|\gg m$ by assumption.%  Furthermore,  
% Also, for low frequency $m=1$ modes $\sbar \simeq -\Omega$, so the
% quantity $|\kappa^2 - \sbar^2| \ll \Omega^2$ in a nearly Keplerian disc
% ($\kappa\sim \Omega$). 
 Hence the RHS of this equation can be neglected. Then 
\begin{align}
  \xi_\phi \simeq -\frac{2\ii\Omega}{\sbar}\xi_R.
\end{align}
For low-frequency $m=1$ modes we have $\sbar \simeq -\Omega$, so
$\xi_\phi\simeq 2\ii\xi_R$, as used in the main text.

### 2014-11-20 paper/discussion.tex
spell checked

\section{Discussion}\label{discussions} 
Having confirmed the $m=1$ spiral instability through numerical
simulations, we now use results from the simulations combined with
linear theory to construct an explanation for their growth. For
simplicitly, we return to 2D discs.

@@

\subsection{Unstable interaction between low-frequency $m=1$ spirals
  and the background disc due to an imposed temperature gradient}
The $m=1$ spiral that emerges in our simulations have low frequency
($\Omega_p < \Omega$),
grows slowly ($\gamma\ll\Omega_p$), and its spatial form is consistent local theory  
(\S\ref{fargo_m1}). Then to a first approximation, we can regard it 
as a neutral, low-frequency $m=1$ mode in satifying the 
dispersion relation Eq. \ref{dispersion} in local theory.  %attribute
                                %growth to temp gradient 
To demonstrate possible instability, %we 
% first construct a neutral low-frequency mode using local theory, 
we show that the torque density acting on such a mode due to the background
temperature gradient is negative, which enforces the mode, because it
has negative angular mometum.  
% Note that
% we will not be assuming a self-gravitating disc until the end of this
% section.

@@

Using $\jlin$ and $T_\mathrm{BG}$, we can obtain a rough estimate of
the growth rate $\gamma$ of linear perturbations due to the background
torque as  
\begin{align}
  2\gamma \sim \frac{T_\mathrm{BG}}{\jlin},
\end{align}
where the factor of two accounts for the fact that the angular momentum
density is quadradtic in the linear perturbations. Inserting the above
expressions for $\jlin$ and $T_\mathrm{BG}$ for $m=1$ gives
\begin{align}
  2\gamma \sim
  -\frac{dc_s^2}{dR}\frac{k}{\Omega}\frac{|\xi_R|^2}{|\xi_R + \ii
    \xi_\phi|^2}. 
%  \lesssim -\frac{dc_s^2}{dR}\frac{k}{\Omega}.
\end{align}
In Appendix \ref{horizontal_displacements}, we argue $\xi_\phi \simeq
2\ii \xi_R$ for the modes of interest. Then 
%To further simplify this expression, we need to relate $\xi_R$ and
%$\xi_\phi$. 
% \begin{align}
%   2\gamma \sim
%   -\frac{dc_s^2}{dR}\frac{1}{k\Sigma^2\Omega}\frac{|\delta\Sigma_1|^2}{|\xi_R
%     + \ii\xi_\phi|^2}.
% \end{align}
% Now, using $\delta\Sigma_m/\Sigma = -\ii k \xi_R - \ii m \xi_\phi/R$
% in the local approximation, we have
% \begin{align}
%   2\gamma \sim -\frac{dc_s^2}{dR}\frac{k}{\Omega}\frac{|\xi_R +
%     \xi_\phi/kR|^2}{|\xi_R + \ii \xi_\phi|^2}
%   \lesssim -\frac{dc_s^2}{dR}\frac{k}{\Omega}.
% \end{align}
% where we used $dc_s^2/dR <0$, $k>0$ and $|kR|\gg 1$ to obtain the
% inequality. 
%where we used $dc_s^2/dR <0$ and $k>0$ to obtain the inequality. 
for the temperature profiles $c_s^2 = c_{s0}^2 (R/R_0)^{-q}$ as
adopted in our disc models,  
\begin{align}
  2\gamma \sim q\frac{c_s^2}{R}\frac{k}{\Omega}. 
\end{align}
For perturbations in a self-gravitating disc with characteristic
wavenumber $k \sim \pi G \Sigma/c_s^2 \sim \Omega/Q c_s$, as seen in
our simulations, we have 
\begin{align}\label{theoretical_rate}
  \gamma \sim \frac{qh}{2Q}\Omega,
\end{align}
where we used $c_s/R\sim h\Omega$. For our fiducial disc models with
$q=1$, $Q=O(1)$, $h=0.05$, this gives $\gamma/\Omega \sim O(10^{-2})$,
consistent with simulation results.

@@

In order to confine an $m=1$ spiral between two $Q$-barriers, we
should have $Q^2(1-\nu^2)=1$ at two radii. Assuming 
Keperlian rotation and a slow pattern speed $\Omega_p\ll\Omega$, this amounts to
\begin{align}
  \left(\frac{R_{Qb}}{R_c}\right)^{-3/2} = 2Q^2(R_{Qb}). 
\end{align}
Then it may be possible to have two $Q$-barriers when the $Q$ profile
of the disc rises more rapidly (decays more slowly) than $R^{-3/2}$
for decreasing (increasing) $R$. One possibility is a gap
opened by a giant planet in a constant-$Q$ disc. In that case $Q$
rises rapidly towards the gap centre since it is a region of low
surface density.  %multi spirals associated with one planet

@@

We expect the external potential to exert a torque on the disc at the
Lindblad and co-rotation resonances. At the outer Lindblad resonance
(OLR), this torque is 
\begin{align}
  \Gamma_L =
  \frac{\pi^2\Sigma_L}{3\Omega_L\Omega_p}
  \left[\left.R_L\frac{d\Phi_\mathrm{ext}}{dR}\right|_L + 2\left(1 - \frac{\Omega_p}{\Omega_L}\right)\Phi_\mathrm{ext}\right]^2,
\end{align}   
where a Keperian disc has been assumed and subscript $L$ denotes
evaluation at the OLR, $R=R_L$. (The inner Lindblad resonance does not
exist for the pattern speeds observed in our simulations.)

>>>>

\section{Discussion}\label{discussions} 
Having confirmed the $m=1$ spiral instability through numerical
simulations, we now use results from the simulations combined with
linear theory to construct an explanation for their growth. For
simplicity, we return to 2D discs.

@@

\subsection{Unstable interaction between low-frequency $m=1$ spirals
  and the background disc due to an imposed temperature gradient}
The $m=1$ spiral that emerges in our simulations have low frequency
($\Omega_p < \Omega$),
grows slowly ($\gamma\ll\Omega_p$), and its spatial form is consistent local theory  
(\S\ref{fargo_m1}). Then to a first approximation, we can regard it 
as a neutral, low-frequency $m=1$ mode in satisfying the 
dispersion relation Eq. \ref{dispersion} in local theory.  %attribute
                                %growth to temp gradient 
To demonstrate possible instability, %we 
% first construct a neutral low-frequency mode using local theory, 
we show that the torque density acting on such a mode due to the background
temperature gradient is negative, which enforces the mode, because it
has negative angular momentum.  
% Note that
% we will not be assuming a self-gravitating disc until the end of this
% section.

@@

Using $\jlin$ and $T_\mathrm{BG}$, we can obtain a rough estimate of
the growth rate $\gamma$ of linear perturbations due to the background
torque as  
\begin{align}
  2\gamma \sim \frac{T_\mathrm{BG}}{\jlin},
\end{align}
where the factor of two accounts for the fact that the angular momentum
density is quadratic in the linear perturbations. Inserting the above
expressions for $\jlin$ and $T_\mathrm{BG}$ for $m=1$ gives
\begin{align}
  2\gamma \sim
  -\frac{dc_s^2}{dR}\frac{k}{\Omega}\frac{|\xi_R|^2}{|\xi_R + \ii
    \xi_\phi|^2}. 
%  \lesssim -\frac{dc_s^2}{dR}\frac{k}{\Omega}.
\end{align}
In Appendix \ref{horizontal_displacements}, we argue $\xi_\phi \simeq
2\ii \xi_R$ for the modes of interest. Then 
%To further simplify this expression, we need to relate $\xi_R$ and
%$\xi_\phi$. 
% \begin{align}
%   2\gamma \sim
%   -\frac{dc_s^2}{dR}\frac{1}{k\Sigma^2\Omega}\frac{|\delta\Sigma_1|^2}{|\xi_R
%     + \ii\xi_\phi|^2}.
% \end{align}
% Now, using $\delta\Sigma_m/\Sigma = -\ii k \xi_R - \ii m \xi_\phi/R$
% in the local approximation, we have
% \begin{align}
%   2\gamma \sim -\frac{dc_s^2}{dR}\frac{k}{\Omega}\frac{|\xi_R +
%     \xi_\phi/kR|^2}{|\xi_R + \ii \xi_\phi|^2}
%   \lesssim -\frac{dc_s^2}{dR}\frac{k}{\Omega}.
% \end{align}
% where we used $dc_s^2/dR <0$, $k>0$ and $|kR|\gg 1$ to obtain the
% inequality. 
%where we used $dc_s^2/dR <0$ and $k>0$ to obtain the inequality. 
for the temperature profiles $c_s^2 = c_{s0}^2 (R/R_0)^{-q}$ as
adopted in our disc models,  
\begin{align}
  2\gamma \sim q\frac{c_s^2}{R}\frac{k}{\Omega}. 
\end{align}
For perturbations in a self-gravitating disc with characteristic
wavenumber $k \sim \pi G \Sigma/c_s^2 \sim \Omega/Q c_s$, as seen in
our simulations, we have 
\begin{align}\label{theoretical_rate}
  \gamma \sim \frac{qh}{2Q}\Omega,
\end{align}
where we used $c_s/R\sim h\Omega$. For our fiducial disc models with
$q=1$, $Q=O(1)$, $h=0.05$, this gives $\gamma/\Omega \sim O(10^{-2})$,
consistent with simulation results.

@@

In order to confine an $m=1$ spiral between two $Q$-barriers, we
should have $Q^2(1-\nu^2)=1$ at two radii. Assuming 
Keplerian rotation and a slow pattern speed $\Omega_p\ll\Omega$, this amounts to
\begin{align}
  \left(\frac{R_{Qb}}{R_c}\right)^{-3/2} = 2Q^2(R_{Qb}). 
\end{align}
Then it may be possible to have two $Q$-barriers when the $Q$ profile
of the disc rises more rapidly (decays more slowly) than $R^{-3/2}$
for decreasing (increasing) $R$. One possibility is a gap
opened by a giant planet in a constant-$Q$ disc. In that case $Q$
rises rapidly towards the gap centre since it is a region of low
surface density.  %multi spirals associated with one planet

@@

We expect the external potential to exert a torque on the disc at the
Lindblad and co-rotation resonances. At the outer Lindblad resonance
(OLR), this torque is 
\begin{align}
  \Gamma_L =
  \frac{\pi^2\Sigma_L}{3\Omega_L\Omega_p}
  \left[\left.R_L\frac{d\Phi_\mathrm{ext}}{dR}\right|_L + 2\left(1 - \frac{\Omega_p}{\Omega_L}\right)\Phi_\mathrm{ext}\right]^2,
\end{align}   
where a Keplerian disc has been assumed and subscript $L$ denotes
evaluation at the OLR, $R=R_L$. (The inner Lindblad resonance does not
exist for the pattern speeds observed in our simulations.)

### 2014-11-20 paper/model.tex
spell checked

\subsection{Disc model and initial conditions}
For the initial disc profile we adopt a modified power-law disc with
surface density given by
\begin{align}
  \Sigma(R) = \Sigma_\mathrm{ref} \left(\frac{R}{R_0}\right)^{-s}\times B(R;
  R_{1}, R_{2}, \epsilon, \delta R), 
\end{align}
where $s$ is the power-law index describing the smooth disc and
$\Sigma_\mathrm{ref}$ is a 
surface density scale chosen by specifying $Q_\mathrm{out}$,
the Keplerian Toomre parameter at $R=R_{2}$,
\begin{align}
  Q_\mathrm{out} = \left.\frac{c_s\Omega_k}{\pi G
    \Sigma}\right|_{R=R_{2}}. 
\end{align}
The bump function
$B(R)$ represents a surface density boost between
$R\in[R_{1},R_{2}]$ by a factor $\epsilon^{-1}>1$,
and $\delta R$ is the transition width between the bump and the
smooth disc. We choose 
\begin{align}
  &B(R) = f_1(R)\times f_2(R),\\
  &f_1(R) = \frac{1}{2}\left(1 - \epsilon\right)\left[1 +
    \tanh\left(\frac{R-R_{1}}{\Delta_1}\right)\right]  + \epsilon,\\
  &f_2(R) = \frac{1}{2}\left(1 - \epsilon\right)\left[1 -
    \tanh\left(\frac{R-R_{2}}{\Delta_2}\right)\right]  + \epsilon,
\end{align}
where $\Delta_{1,2} = \delta R \times H(R_{1,2})$. The bump function
mimicks mass built-up in $R\in[R_1,R_2]$.

@@

\subsection{FARGO}
Our primary code is FARGO with self-gravity \citep{baruteau08}. This
is a popular, simple finite-difference code for 2D discs. `FARGO' refers
to its azimuthal transport algorithm, which circumvents time-step
contraint set by the inner disc boundary \citep{masset00a,masset00b}.

The 2D disc occupies
$R\in[R_\mathrm{min},R_\mathrm{max}],\,\phi\in[0,2\pi]$ and is 
divided into $(N_R,N_\phi)$ grids, logarithimically spaced in radius and
uniformly spaced in azimuth. At radial boundaries we set the
hydrodynamic variables to their initial values.

The 2D Poisson equation is solved in integral form, 
\begin{align}\label{2d_grav}
  &\Phi_{d,z=0}(R,\phi) \notag \\
  &=-\int_{R_\mathrm{min}}^{R_\mathrm{max}} \int_0^{2\pi}
  \frac{G\Sigma(R^\prime,\phi^\prime)R^\prime dR^\prime d\phi^\prime}{\sqrt{R^2+R^{\prime 2} -
      2RR^\prime\cos{(\phi - \phi^\prime)} + \epsilon_g^2}}, 
\end{align}
using Fast Fouier Transform (FFT), where $\epsilon_g$ is a softening
length to prevent a numerical singularity. The FFT approach requires
$\epsilon_g\propto R$ \citep{baruteau08}. In FARGO, $\epsilon_g$ is
set to a fraction of $hR$.

@@

%used in lin 2012 
\subsection{PLUTO} 
PLUTO is a general-purpose Godunov code \citep{mignone07}. The grid
setup is the same as that adopted in ZEUS-MP above. We configure the
code similarly to that used in \cite{lin14}: piecewise linear
reconstruction, a Roe solver and second order Runge-Kutta time
integration. We also enable the FARGO algorithm for azimuthal
transport.

@@

In a linear analysis, one assumes a steady axisymmetric background state, such as
that defined in \S\ref{model}, which is then perturbed such that
\begin{align}  
  \Sigma \to \Sigma(R) + \delta\Sigma_m(R)\exp{\left[\ii\left(-\sigma t +
        m\phi\right)\right]}, 
\end{align}
and similarly for other variables. We take $m>0$. Since the background is
axisymmetric, we identify $\delta\Sigma_m(R)\exp{\left(-\ii\sigma
    t\right)} = \Sigma_m$ as defined for diagnostics. 
The linearised mass and momentum equations are
\begin{align}
  &-\ii\sbar \delta\Sigma_m = -\frac{1}{R}\frac{d}{dR}\left(R\Sigma\delta
    v_{Rm}\right) - \frac{\ii m \Sigma}{R}\delta v_{\phi m}, \\
  &-\ii\sbar\delta v_{Rm} - 2\Omega \delta v_{\phi m} = -
  c_s^2(R)\frac{d}{dR}\left(\frac{\delta\Sigma_m}{\Sigma}\right) - \frac{d}{dR}\delta\Phi_m,\\
  & -\ii\sbar\delta v_{\phi m} + \frac{\kappa^2}{2\Omega}\delta v_{Rm} =
  -\frac{\ii m }{R}\left(c_s^2\frac{\delta\Sigma_m}{\Sigma} + \delta\Phi_m\right),
\end{align}
where $\sbar = \sigma - m\Omega$ and $\kappa^2 =
R^{-3}\p_R(R^4\Omega^2)$ is the square of the epicycle frequency. A
locally isothermal equation of state has been assumed.    
The linearised Poisson equation is 
\begin{align}
  \nabla^2\delta\Phi_m = 4\pi G \delta\Sigma_m \delta(z). 
\end{align}

@@

\subsection{Global angular momentum conservation for linear
  perturbations} 
It can be shown that linear perturbations with $\phi$-dependence in the form
$\exp{(\ii m\phi)}$ satisfies angular momentum conservation
in the form 
\begin{align} 
  \frac{\p \jlin}{\p t} + \nabla\cdot\bm{F} = \tbg, 
\end{align}
\citep[e.g.][]{narayan87,ryu92,lin93b} where
\begin{align}\label{lin_ang_mom_cons}
  \jlin \equiv
  -\frac{m\Sigma}{2}\imag\left(\bm{\xi}^*\cdot\frac{\p\bm{\xi}}{\p
      t} + \Omega \hat{\bm{k}}\cdot\bm{\xi}\times\bm{\xi}^* + \ii
    m \Omega |\bm{\xi}|^2\right) 
\end{align}
is the angular momentum density, $\bm{\xi}$ is the Lagragian
displacement, and $\bm{F}$ is the vertically-integrated angular
momentum flux consisting of a Reynolds stress and a gravitational
stress \citep{lin11b}. Explicit expressions for $\bm{\xi}$ can be found in,
e.g. \cite{papaloizou85}. 
%and $\bm{F}$ can be read off Eq. \ref{angflux1}.

@@

The background torque density $\tbg$ is 
\begin{align}\label{baroclinic_torque}
  \tbg \equiv
  -\frac{m}{2}\imag\left(\delta\Sigma_m\xi_R^*\frac{dc_s^2}{dR}\right), 
\end{align}
and arises because we have adopted a locally isothermal equation of
state. In a barotropic fluid, such as a strictly isothermal disc,
$\tbg$ vanishes and the total angular momentum associated with
the perturbation is conserved, provided that there is no net angular
momeumtum flux.% and the perturbations are regular.

@@

A detailed discussion of the properties of local density waves 
is given in \cite{shu91}. An important result is that
waves interior to co-rotation ($R<R_c$) have negative angular momentum, while
waves outside co-rotation ($R>R_c$) have positive angular
momentum. The local approximation becomes less good for gobal modes,
but will nevertheless be useful for results interpretation.

>>>>

\subsection{Disc model and initial conditions}
For the initial disc profile we adopt a modified power-law disc with
surface density given by
\begin{align}
  \Sigma(R) = \Sigma_\mathrm{ref} \left(\frac{R}{R_0}\right)^{-s}\times B(R;
  R_{1}, R_{2}, \epsilon, \delta R), 
\end{align}
where $s$ is the power-law index describing the smooth disc and
$\Sigma_\mathrm{ref}$ is a 
surface density scale chosen by specifying $Q_\mathrm{out}$,
the Keplerian Toomre parameter at $R=R_{2}$,
\begin{align}
  Q_\mathrm{out} = \left.\frac{c_s\Omega_k}{\pi G
    \Sigma}\right|_{R=R_{2}}. 
\end{align}
The bump function
$B(R)$ represents a surface density boost between
$R\in[R_{1},R_{2}]$ by a factor $\epsilon^{-1}>1$,
and $\delta R$ is the transition width between the bump and the
smooth disc. We choose 
\begin{align}
  &B(R) = f_1(R)\times f_2(R),\\
  &f_1(R) = \frac{1}{2}\left(1 - \epsilon\right)\left[1 +
    \tanh\left(\frac{R-R_{1}}{\Delta_1}\right)\right]  + \epsilon,\\
  &f_2(R) = \frac{1}{2}\left(1 - \epsilon\right)\left[1 -
    \tanh\left(\frac{R-R_{2}}{\Delta_2}\right)\right]  + \epsilon,
\end{align}
where $\Delta_{1,2} = \delta R \times H(R_{1,2})$. The bump function
mimics mass built-up in $R\in[R_1,R_2]$.

@@

\subsection{FARGO}
Our primary code is FARGO with self-gravity \citep{baruteau08}. This
is a popular, simple finite-difference code for 2D discs. `FARGO' refers
to its azimuthal transport algorithm, which circumvents time-step
constraint set by the inner disc boundary \citep{masset00a,masset00b}.

The 2D disc occupies
$R\in[R_\mathrm{min},R_\mathrm{max}],\,\phi\in[0,2\pi]$ and is 
divided into $(N_R,N_\phi)$ grids, logarithmically spaced in radius and
uniformly spaced in azimuth. At radial boundaries we set the
hydrodynamic variables to their initial values.

The 2D Poisson equation is solved in integral form, 
\begin{align}\label{2d_grav}
  &\Phi_{d,z=0}(R,\phi) \notag \\
  &=-\int_{R_\mathrm{min}}^{R_\mathrm{max}} \int_0^{2\pi}
  \frac{G\Sigma(R^\prime,\phi^\prime)R^\prime dR^\prime d\phi^\prime}{\sqrt{R^2+R^{\prime 2} -
      2RR^\prime\cos{(\phi - \phi^\prime)} + \epsilon_g^2}}, 
\end{align}
using Fast Fourier Transform (FFT), where $\epsilon_g$ is a softening
length to prevent a numerical singularity. The FFT approach requires
$\epsilon_g\propto R$ \citep{baruteau08}. In FARGO, $\epsilon_g$ is
set to a fraction of $hR$.

@@

%used in lin 2012 
\subsection{PLUTO} 
PLUTO is a general-purpose Godunov code \citep{mignone07}. The grid
setup is the same as that adopted in ZEUS-MP above. We configure the
code similarly to that used in \cite{lin14}: piece-wise linear
reconstruction, a Roe solver and second order Runge-Kutta time
integration. We also enable the FARGO algorithm for azimuthal
transport.

@@

In a linear analysis, one assumes a steady axisymmetric background state, such as
that defined in \S\ref{model}, which is then perturbed such that
\begin{align}  
  \Sigma \to \Sigma(R) + \delta\Sigma_m(R)\exp{\left[\ii\left(-\sigma t +
        m\phi\right)\right]}, 
\end{align}
and similarly for other variables. We take $m>0$. Since the background is
axisymmetric, we identify $\delta\Sigma_m(R)\exp{\left(-\ii\sigma
    t\right)} = \Sigma_m$ as defined for diagnostics. 
The linearised mass and momentum equations are
\begin{align}
  &-\ii\sbar \delta\Sigma_m = -\frac{1}{R}\frac{d}{dR}\left(R\Sigma\delta
    v_{Rm}\right) - \frac{\ii m \Sigma}{R}\delta v_{\phi m}, \\
  &-\ii\sbar\delta v_{Rm} - 2\Omega \delta v_{\phi m} = -
  c_s^2(R)\frac{d}{dR}\left(\frac{\delta\Sigma_m}{\Sigma}\right) - \frac{d}{dR}\delta\Phi_m,\\
  & -\ii\sbar\delta v_{\phi m} + \frac{\kappa^2}{2\Omega}\delta v_{Rm} =
  -\frac{\ii m }{R}\left(c_s^2\frac{\delta\Sigma_m}{\Sigma} + \delta\Phi_m\right),
\end{align}
where $\sbar = \sigma - m\Omega$ and $\kappa^2 =
R^{-3}\p_R(R^4\Omega^2)$ is the square of the epicyclic frequency. A
locally isothermal equation of state has been assumed.    
The linearised Poisson equation is 
\begin{align}
  \nabla^2\delta\Phi_m = 4\pi G \delta\Sigma_m \delta(z). 
\end{align}

@@

\subsection{Global angular momentum conservation for linear
  perturbations} 
It can be shown that linear perturbations with $\phi$-dependence in the form
$\exp{(\ii m\phi)}$ satisfies angular momentum conservation
in the form 
\begin{align} 
  \frac{\p \jlin}{\p t} + \nabla\cdot\bm{F} = \tbg, 
\end{align}
\citep[e.g.][]{narayan87,ryu92,lin93b} where
\begin{align}\label{lin_ang_mom_cons}
  \jlin \equiv
  -\frac{m\Sigma}{2}\imag\left(\bm{\xi}^*\cdot\frac{\p\bm{\xi}}{\p
      t} + \Omega \hat{\bm{k}}\cdot\bm{\xi}\times\bm{\xi}^* + \ii
    m \Omega |\bm{\xi}|^2\right) 
\end{align}
is the angular momentum density, $\bm{\xi}$ is the Lagrangian
displacement, and $\bm{F}$ is the vertically-integrated angular
momentum flux consisting of a Reynolds stress and a gravitational
stress \citep{lin11b}. Explicit expressions for $\bm{\xi}$ can be found in,
e.g. \cite{papaloizou85}. 
%and $\bm{F}$ can be read off Eq. \ref{angflux1}.

@@

The background torque density $\tbg$ is 
\begin{align}\label{baroclinic_torque}
  \tbg \equiv
  -\frac{m}{2}\imag\left(\delta\Sigma_m\xi_R^*\frac{dc_s^2}{dR}\right), 
\end{align}
and arises because we have adopted a locally isothermal equation of
state. In a barotropic fluid, such as a strictly isothermal disc,
$\tbg$ vanishes and the total angular momentum associated with
the perturbation is conserved, provided that there is no net angular
momentum flux.% and the perturbations are regular.

@@

A detailed discussion of the properties of local density waves 
is given in \cite{shu91}. An important result is that
waves interior to co-rotation ($R<R_c$) have negative angular momentum, while
waves outside co-rotation ($R>R_c$) have positive angular
momentum. The local approximation becomes less good for global modes,
but will nevertheless be useful for results interpretation.
